%
% 6.006 problem set 7E
%
\documentclass[12pt,twoside]{article}

\input{macros}

\usepackage{amsmath}
\usepackage{url}
\usepackage{mdwlist}
\usepackage{graphicx}
\usepackage{clrscode3e}
\newcommand{\isnotequal}{\mathrel{\scalebox{0.8}[1]{!}\hspace*{1pt}\scalebox{0.8}[1]{=}}}
\usepackage{listings}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{matrix}
\usetikzlibrary{positioning}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{shapes.misc}
\usetikzlibrary{trees}

\setlength{\oddsidemargin}{0pt}
\setlength{\evensidemargin}{0pt}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{0in}
\setlength{\textheight}{8.5in}

% Fill these in!
\newcommand{\theproblemsetnum}{7E}
\newcommand{\releasedate}{December 7, 2011}
\newcommand{\partaduedate}{Thursday, December 8}
\newcommand{\tabUnit}{3ex}
\newcommand{\tabT}{\hspace*{\tabUnit}}

\begin{document}

\handout{Problem Set \theproblemsetnum}{\releasedate}
You will need to submit the \verb|ps7e_answer.tex| file containing your grading
explanation to the Stellar assignment ``Problem Set 7E -- LaTeX Template'' by
\textbf{Thursday, December 8th, 11:59PM}. Your grade for Problem 2(t) in
Problem Set 7 will be based on both your solutions and the grading explanation.
\medskip

\hrulefill

\begin{problems}

\problem Please write a short explanation of your pseudo-code for Problem 2(t)
in Problem Set 7. Assume the grader is familiar with the problem and with the
staff solution. If your solution is very similar to the official solution,
please state that, and point out and explain any differences. If your solution
is correct, but very different from the official solution, write a brief
explanation to convince the grader that your solution is correct. If your
solution is partially correct, point out the parts that you got right, and
explain your mistakes.

\textbf{Answer:}

%%% PROBLEM 2(tE) ANSWER START %%%
Clearly the runtime I provided of $\Theta(cash * count * total)$ is incorrect because I meant cash to be total amount of cash possible, which is $total$. If I had written this correctly, then I would have obtained the correct runtime of $\Theta(count * total^2)$. 

My algorithm and recurrence are mostly correct, although I missed a couple of edge cases. My code basically implemented the second solution, but with an extra maximum. Although my recurrence did not contain it, my pseudocode set $profit[c,s,0] = cash$, which is the first condition in solution 2's recurrence. I did forget to set $profit[c,s,x] = profit[c,s-1, limit[s-1]]$ if $x = 0$ or $start[s] > c$. 

The main part of the recurrence was mostly the same, except I took a maximum $max_{i \in \{0, \ldots, limit[s-1] \} }$ over $profit[c,s-1,i]$. This should give a correct answer, because instead of setting the last argument of $profit[c,s,x]$ as the new limit like in the solutions, I set it as the number of stocks bought. Thus, taking the maximum over all the possible $x$'s of buying $x$ of a given stock will be equivalent to the solution's answer of taking the maximum profit to buy a stock with a given limit. This is because $x$ is bounded by the same limit as in the solution's answer. I included the best profit for each stock in a new table $bestprofit$, which would be equivalent to the $profit$ table contained in the first solution.
%%% PROBLEM 2(tE) ANSWER END %%%

\end{problems}

\end{document}
